Approximately Exact Calculations for Linear Mixed Models

### 2015-05-12 ms.tex
clarified Eq 3, 4, 5

As \cite{hodges:2013} further observes, if $\beta$ is given an improper flat prior and $\sigesq$ and $\sigssq$ are given conjugate priors --- say $\sigesq \sim \IG(\alpha_e,\beta_e)$ and $\sigssq \sim \IG(\alpha_s,\beta_s)$ --- then \begin{equation*}
  -(\alpha_e+1) \log\sigesq - \beta_e/\sigesq -(\alpha_s+1) \log\sigssq - \beta_s/\sigssq
\end{equation*}
is added to \eqref{eq:reexpress} to yield
the log posterior
\begin{equation}
\label{eq:logpost}
  \begin{split}
  \log\pi(\sigesq,\sigssq\g y) &=
  B - \frac{n_e + 2\alpha_e + 2}{2}\log(\sigesq) -
    \frac{y^t \Gamma_c \Gamma^t_c y + 2\beta_e}{2\sigesq}\\
    &\qquad\qquad - \frac{2\alpha_s + 2} {2}\log(\sigssq) - \frac{2\beta_s}{2\sigssq}
    - \frac{1}{2} \sum_{j=1}^{s_z} \left[ \log(\ass) + \frac{\hat v_j^2}{\ass}\right]\\
    &= B - \frac{1}{2}\sum_j\left[ c_j \log(\abss) + \frac{d_j}{\abss}\right]
  \end{split}
\end{equation}
for known nonnegative constants $\{a_j, b_j, c_j, d_j\}$ ---
still a linear combination of logs and inverses of linear functions $\abss$.  Our derivation will proceed from \eqref{eq:logpost}, though it applies also to \eqref{eq:reexpress} and other linear combinations of logs and inverses of linear functions $\abss$.  We use $f$ to denote such functions generically.

>>>>

As \cite{hodges:2013} further observes, if $\beta$ is given an improper flat prior and $\sigesq$ and $\sigssq$ are given conjugate priors --- say $\sigesq \sim \IG(\alpha_e,\beta_e)$ and $\sigssq \sim \IG(\alpha_s,\beta_s)$ --- then \begin{equation*}
  -(\alpha_e+1) \log\sigesq - \beta_e/\sigesq -(\alpha_s+1) \log\sigssq - \beta_s/\sigssq
\end{equation*}
is added to \eqref{eq:reexpress} to yield
the log posterior
\begin{equation}
\label{eq:logpost}
  \begin{split}
  \log\pi(\sigesq,\sigssq\g y) &=
  B - \frac{n_e + 2\alpha_e + 2}{2}\log(\sigesq) -
    \frac{y^t \Gamma_c \Gamma^t_c y + 2\beta_e}{2\sigesq}\\
    &\qquad\qquad - \frac{2\alpha_s + 2} {2}\log(\sigssq) - \frac{2\beta_s}{2\sigssq}
    - \frac{1}{2} \sum_{j=1}^{s_z} \left[ \log(\ass) + \frac{\hat v_j^2}{\ass}\right].
  \end{split}
\end{equation}
Equations~\eqref{eq:reexpress} and~\eqref{eq:logpost} can both be written as in \eqref{eq:sum}: a summation of multiples of logs and inverses of linear combinations $\abss$, where the summands with $j=s_Z+1$ and $j=s_Z+2$ are for the terms involving only $\sigesq$ and $\sigssq$, respectively.
\begin{equation}
\label{eq:sum}
  -\frac{1}{2} \sum_{j=1}^{s_Z+2}\left[ c_j \log(\abss) + \frac{d_j}{\abss}\right]
\end{equation}
where
\begin{itemize}[label=---]
\item for $j = 1, \dots, s_Z$,
  \begin{align*}
    a_j &> 0\ \text{and is derived in the Appendix}\\
    b_j &=1\\
    c_j &=1\\
    d_j &= \hat v_j^2\ \text{and is derived in the Appendix}
  \end{align*}
\item for $j = s_Z +1$,
  \begin{align*}
    a_j &= 0\\
    b_j &= 1\\
    c_j &= \begin{cases}
                 n_e & \text{for $\logRLorig$ in \eqref{eq:reexpress}}\\
                 n_e + 2\alpha_e + 2 & \text{for the $\log$ posterior in \eqref{eq:logpost}}
              \end{cases}\\
    d_j &= \begin{cases}
                 \frac{y^t \Gamma_c \Gamma^t_c y}{\sigesq} & \text{for $\logRLorig$ in \eqref{eq:reexpress}}\\
                 \frac{y^t \Gamma_c \Gamma^t_c y + 2\beta_e}{\sigesq} & \text{for the $\log$ posterior in \eqref{eq:logpost}}
              \end{cases}
  \end{align*}
\item for $j = s_Z+2$,
  \begin{align*}
    a_j &= 1\\
    b_j &= 0\\
    c_j &= \begin{cases}
      0 & \text{for $\logRLorig$ in \eqref{eq:reexpress}}\\
      2 \alpha_s + 2 & \text{for the $\log$ posterior in \eqref{eq:logpost}}
    \end{cases}\\
    d_j &= \begin{cases}
      0 & \text{for $\logRLorig$ in \eqref{eq:reexpress}}\\
      2 \beta_s & \text{for the $\log$ posterior in \eqref{eq:logpost}}.
    \end{cases}
  \end{align*}
\end{itemize}
Our derivation proceeds from \eqref{eq:sum}.  We use $f$ to denote the target function generically, either~\eqref{eq:reexpress} or~\eqref{eq:logpost}.
